--- 
title_meta  : Chapter 3
title       : WEEK 3, Linear regression (deadline 28.11. 23:59)
description : "This weeks exercises will focus on linear regression. You will form a few models yourself and train to interpret the results from those models. You will also learn to evaluate the assumptions of the models."
 
--- type:NormalExercise lang:r xp:100 skills:1 key:3d083aaefb
## Linear regression with one explanatory variable
 
In this exercise you will form a linear regression model, see the results and make some interpretations based on those. In R linear regression can be done using lm()- function, for example: lm(cardiac$dobEF~cardiac$baseEF). Before the ~ sign you can put only one variable, the outcome variable. After it you can put one or multiple explanatory variables devided by the + sign. 
 
*** =instructions
 - Form a linear regression model, in which you will have systolic blood pressure (sbp) as an outcome variable and age as an explanatory variable. Use the Cardiac-data. (Don't use the with-function this time.)

 - Interpret the results: 
   + What is the regression coefficient?
   + What is the intercept?
   + What is the variance explained?
 
*** =hint
Use formula lm(data$outcome~data$explanatory). To show the results, use summary-fucntion. 
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# Form the model:

# Check the results:

# Fill the following sentences with the right number:
# Regression coefficient is:
# Intercept is:
# Variance explained is:

```
 
*** =solution
```{r}
reg1<-lm(cardiac$sbp~cardiac$age)
summary(reg1)
# Fill the following sentences with the right number:
# Regression coefficient is: -0.2203
# Intercept is: 161.7502
# Variance explained is: 0.00349
```
 
*** =sct
```{r}
test_student_typed("lm(cardiac$sbp~cardiac$age)", not_typed_msg = "Make sure you have put the variables in the right order, and they are divided by the ~ sign.")
test_student_typed("summary(", not_typed_msg = "Make sure you have used the summary-function to see the results.")
test_student_typed("Regression coefficient is: -0.22", not_typed_msg = "Make sure you have typed the right regression coefficient.")
test_student_typed("Intercept is: 161.7", not_typed_msg = "Make sure you have typed the right intercept.")
test_student_typed("Variance explained is: 0.00349", not_typed_msg = "Make sure you have typed the right variance explained.")


```

--- type:NormalExercise lang:r xp:100 skills:1 key:harj32
## Linear regression with multiple explanatory variables
 
In the previous exercise you made a linear regression using one explanatory variables. In this exercise you will add explanatory variables. 
 
*** =instructions
 - Form a linear regression model, in which you will have systolic blood pressure (sbp) as an outcome variable and age, sex and ejection fraction (baseEF) as an explanatory variable. Use the Cardiac-data. (Don't use the with-function this time.)
 Interpret the results: 
   - What is the regression coefficient for age?
   - What is the regression coefficient for gender?
   - What is the regression coefficient for ejection fraction?
   - What is the intercept?
   - What is the variance explained?
 
*** =hint
Use formula lm(data$outcome~data$explanatory1+data$explanatory2+ etc).  
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# Form the model

# Check the results

# Fill the following sentences with the right numbers:
# Regression coefficient for age is: 
# Regression coefficient for gender is: 
# Regression coefficient for ejection fraction is: 
# Intercept is: 
# Variance explained is: 
 
```
 
*** =solution
```{r}
reg<-lm(cardiac$sbp~cardiac$age+cardiac$gender+cardiac$baseEF)
summary(reg)
# Regression coefficient for age is: -0.2219

# Regression coefficient for gender is: -0.1973

# Regression coefficient for ejection fraction is: 0.4314

# Intercept is: 137.9916

# Variance explained is: 0.001469

```
 
*** =sct
```{r}
test_student_typed("lm(cardiac$sbp~cardiac$age+cardiac$gender+cardiac$baseEF)", not_typed_msg = "Make sure you have put the variables in the right order, and they are divided by the + sign.")
test_student_typed("summary(", not_typed_msg = "Make sure you have used the summary-function to see the results.")
test_student_typed("Regression coefficient for age is: -0.22", not_typed_msg = "Make sure you have typed the right regression coefficient for age.")
test_student_typed("Regression coefficient for gender is: -0.1973", not_typed_msg = "Make sure you have typed the right regression coefficient for gender.")
test_student_typed("Regression coefficient for ejection fraction is: 0.43", not_typed_msg = "Make sure you have typed the right regression coefficient for ejection fraction.")
test_student_typed("Intercept is: 137.9", not_typed_msg = "Make sure you have typed the right intercept.")
test_student_typed("Variance explained is: 0.01", not_typed_msg = "Make sure you have typed the right variance explained.")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:99950f38b0
## Checking the assumptions for your model graphically
 
Now you know how to form linear regression models and check their results. In this exercise you will learn how to check how your model fits the assumptions (normality). You can do this for example creating a Q-Q plot from your model using qqnorm(studred(model))- command, or plotting a histogram from the residuals of your model typing hist(studres(model)).
 
*** =instructions
The models you created in the previous exercises are named reg1 and reg2, accordingl with the following commands:
     -reg1<-lm(cardiac$sbp~cardiac$age)
     -reg2<-lm(cardiac$sbp~cardiac$age+cardiac$gender+cardiac$baseEF)
You don't need to run these by yourself anymore!
 - Create Q-Q plot from your 'one explanatory variable' model, reg1
 - Create Q-Q plot from your 'multiple explanatory variables' model, reg2
*** =hint
  
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
library(MASS)

reg1<-lm(cardiac$sbp~cardiac$age)
reg2<-lm(cardiac$sbp~cardiac$age+cardiac$gender+cardiac$baseEF)
```
 
*** =sample_code
```{r}
## Create Q-Q plot from your 'one explanatory variable' model, reg1 

## Plot histogram from that same model's residuals

## Create Q-Q plot from your 'multiple explanatory variables' model, reg2 

## Plot histogram from that same model's residuals

## Now plot histogram from your outcome- variable, cardiac$sbp, and compare it with the histograms of residuals. What can you see?
 
```
 
*** =solution
```{r}
qqnorm(studred(reg1))
hist(studres(reg1))
qqnorm(studred(reg2))
hist(studres(reg2))
hist(cardiac$sbp)
```
 
*** =sct
```{r}

```
