--- 
title_meta  : Chapter 2
title       : Plots
description : "In this chapter, you will learn to plot."
 
--- type:NormalExercise lang:r xp:100 skills:1 key:exercise0_1
## THE WARM UP STARTS HERE: How does DataCamp work
 
In the box below you can see the instructions for the exercise. 
On the right in the upper window you can type your code. That window is called editor. 
You can execute the code by pressing Ctrl+Enter (Cmd+Enter in Mac) when the cursor is on the same line with your command,
and when you are done with the exercises, you can press the Submit-button.
If your command is on multiple lines you must select all those rows and then Ctrl+Enter or Cmd+Enter.
When you run your commands, the results appear in the window under the editor. That window is called console.

Tips for iPad users:
In case you are using an iPad, you can just go ahead and run all the commands by pressing the Submit-button. 
If you do this when you are not yet 
finished and some sort of error message appears, you can try to comment out the parts that you have not completed yet. 
That can be done by inserting a hashtag `#` at the beginning of the not completed lines.
 
You can use R as a calculator. You can run basic mathematical operations like this: 
 
- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`
 
In order to understand better your own code, or in case you want to save it as an actual R script 
(It is possible as a simple copy-paste if you have R installed in your computer.) it is good to comment on your code. 
At the beginning when the code is still not very clear to you, it helps you to understand what the code does. When the time
goes by, these weird codewords start to mean something to you and it becomes easier, but the comments still help you
and are especially useful if you need to share your code with another person.
Commenting can be done by typing `#` at the beginning of the row. Any text you type on that row will be considered as a comment. 
For example _Write your name here_ in the editor window is a comment.

Don't be afraid of making mistakes, you do not loose any points by submitting an incorrect solution. 
If your solution is not correct, you will receive an error message and you can simply try again.
If, on the other hand you decide to take a hint, you will loose some points and if you decide to look at the ready-made 
solution, you will loose your points for that exercise.
 
*** =instructions
 - These first exercises are a warm up to get to know how does this environment work.
 - Whenever the commented line asks you to do something, you should write the answer under that line unless told otherwise.

*** =hint
Just follow the instructions written in the editor. If you do exactly as they say, you should be able to complete this
exercise. Try submitting your code and maybe the error messages help you to continue. If this doesn't help, ask the course
staff or a friend or try posting your question on Moodle, maybe there are other students struggling with the same problem!
 
*** =pre_exercise_code
```{r}
 # no pec
```
 
*** =sample_code
```{r}
# Write your name here

 
# Write here numbers from one to ten without spaces

 
# Calculate the sum of the following numbers: 1,3,5,7,9

 
# If Matti has 100 apples and he gives Mari 35 of them, how much apples does Matti have left? Calculate.
 

```
 
*** =solution
```{r}
# Write your name here
Maija Meikalainen
 
# Write here numbers from one to ten without spaces
12345678910
 
# Calculate the sum of the following numbers: 1,3,5,7,9
1+3+5+7+9
 
# If Matti has 100 apples and he gives Mari 35 of them, how much apples does Matti have left? Calculate.
100-35
```
 
*** =sct
```{r}
test_output_contains("12345678910", incorrect_msg = "When you are writing the numbers from one to ten, make sure there are no spaces between them.")
test_output_contains("25", incorrect_msg = "Your sum is not quite correct, try again.")
test_output_contains("65", incorrect_msg = "Try counting the apples again.")
success_msg("It seems that you are getting the idea. Let's still continue the warm up for a little bit.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:exercise0_2
## Warming up your special characters
In the previous exercise we learned how to write simple things on the editor and what they look like in the console.
This time we want to focus on special characters so that when they come up in the actual exercises you already know
a little bit about them.

*** =instructions

Calculate the mean and standard deviation for that same data set as in previous exercise using function mean() and sd(). You should get the same results as in previous exercise. As a reminder: the ejectionfractions (in percents) were as follow: : 64, 57, 52, 37, 57, 62, 60, 57, 57 and 37.

*** =hint
Make sure you have not forget to put those numbers in vector c()!
 
*** =pre_exercise_code
```{r}
 # no pec
```
 
*** =sample_code
```{r}
# Calculate the arithmetic mean of these numbers: 1, 3, 8, 10, 12
mean(c(1, 3, 8, 10, 12))
 
# Calculate the standard deviation of these numbers: 1, 3, 8, 10, 12
sd(c(1, 3, 8, 10, 12))
 
# Calculate the arithmetic mean of the ejectiofraction as shown in example above

 
# Calculate the standard deviation of the ejectiofraction as shown in example above
 

```
 
*** =solution
```{r}
# Calculate the arithmetic mean of these numbers: 1, 3, 8, 10, 12
mean(c(1, 3, 8, 10, 12))
 
# Calculate the standard deviation of these numbers: 1, 3, 8, 10, 12
sd(c(1, 3, 8, 10, 12))

# Calculate the arithmetic mean of the ejectiofraction
mean(c(64,57,52,37,57,62,60,57, 57 ,37))
 
# Calculate the standard deviation of the ejectiofraction
sd(c(64,57,52,37,57,62,60,57, 57 ,37))
```
 
*** =sct
```{r}
test_output_contains("54", incorrect_msg = "Make sure that you have put the correct values inside mean()-function.")
test_output_contains("9.533", incorrect_msg = "Make sure that you have put the correct values inside sd()-function.")
test_student_typed("mean(", not_typed_msg = "Make sure you used mean()- function.")
test_student_typed("sd(", not_typed_msg = "Make sure you used sd()- function.")
success_msg("Super! Now let's move on to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:exercise0_3
## Checking out the data! 
 
When you get a new data set, it is a good practice to first give it a quick overview: to check how many samples you have, how many variables you have, how many females / males, and what is the age distribution, for example. To get the number of samples and variables in your data, just type dim(data). For categorical variables, table() is a good way, and for continous variables use summary() to get an overview of that specific variable. If you need to know what variables you have in your data, just type colnames(data).
 
*** =instructions
 - Find out how many observations does the cardiac-data contain. 
 - Find out how many observations does the diabetes-data contain.
 - Find out how many females are in cardiac-data.
 - Find out how many female are in diabetes- data.
 - What is the age-range in cardiac-data?
 - What is the age-range in diabetes-data?

*** =hint
Checking the dimensions of the dataset is a good way to find out how many observations are there. 
Remember to connect variables to the dataset with a '$' sign. Remember also the table() and summary()- functions!
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# Hint! You can refer to our data sets just by typing "cardiac" and "diabetes". To refer to any variable in that data just type it in the form data$variable, for example cardiac$gender.

# Fill in the sample sizes:
# Cardiac:
# Diabetes: 

# Females in cardiac-data:
# Females in diabetes-data: 
 
# Check age distribution in cardiac-data (summary of the variable age in cardiac-data):


# Chekc age distribution in diabetes-data (summary of the variable age in diabetes-data):
 
 
```
 
*** =solution
```{r}
# Fill in the sample sizes:
# Cardiac: 558
# Diabetes: 403

# Females in cardiac-data: 338
# Females in diabetes-data: 234
 
# Check age distribution in cardiac-data (summary of the variable 'age' in cardiac-data):
summary(cardiac$age)

# Chekc age distribution in diabetes-data (summary of the variable 'age' in diabetes-data):
summary(diabetes$age)
 
```
 
*** =sct
```{r}
test_student_typed("558", not_typed_msg = "Make sure that you have the right sample size for cardiac-data.")
test_student_typed("403", not_typed_msg = "Make sure that you have the right sample size for diabetes-data.")
test_student_typed("338", not_typed_msg = "Make sure that you have the right female number for cardiac-data.")
test_student_typed("234", not_typed_msg = "Make sure that you have the right female number for diabetes-data.")
test_student_typed("summary(cardiac$age)", not_typed_msg = "Make sure you checked the summary of the variable age in cardiac- dataset correctly.")
test_student_typed("summary(diabetes$age)", not_typed_msg = "Make sure you checked the summary of the variable age in diabetes- dataset correctly.")

success_msg("Great! You're such a master at R! ")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:exercise0_4
## Basic plotting
 
When checking out your data, the good next step is to visualize the variables. This is how you get the first idea of their distributions and can spot the possible outliers (the kind of values that are very different from the other values.) Propably most commonly used plotting commands are hist() for histrogram, barplot() for barplot, plot() for scatter plot and boxplot() for boxplot. 
 
*** =instructions
 - Draw a barplot of the variable 'gender' in diabetes-dataset.
 
 - Draw a histogram of the variable 'age' in cardiac-dataset.
 
 - Draw a histogram of the variable 'bhr' (= "basal heart rate") in cardiac-dataset. Is there anything weird?
 
 - Draw a boxplot for cholesterol ('chol') in diabetes- data by gender.
 
 - Draw a scatter-plot for the variables maximum heart rate ('maxhr') and systolic blood pressure ('sbp') in cardiac-data
 
 - Check also the correlation of maximum heart rate ('maxhr') and systolic blood pressure ('sbp') in cardiac-data.


*** =hint
Make sure you have the right variable(s) in your plotting- command, and you have used the right plotting- command! Make also sure you have not made any spelling mistakes.
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# Example1: Draw a barplot of the variable 'gender' in cardiac-dataset
 barplot(table(cardiac$gender))
 
# Example2: Draw a histogram of the variable 'cholmmol' (cholesterol in mmol) in diabetes-dataset
 hist(diabetes$cholmmol)
 
# Draw a barplot of the variable 'gender' in diabetes-dataset

 
# Draw a histogram of the variable 'age' in cardiac-dataset


# Draw a histogram of the variable 'bhr' (="basal heart rate") in cardiac-dataset. Is there anything weird?

 
# Draw a boxplot for cholesterol ('chol') in diabetes- data by gender.

# Draw a scatter-plot for the variables maximum heart rate ('maxhr') and systolic blood pressure ('sbp') in cardiac-data 

# Check the correlation of maximum heart rate ('maxhr') and systolic blood pressure ('sbp') in cardiac-data.

 
```
 
*** =solution
```{r}
# Example1: Draw a barplot of the variable 'gender' in cardiac-dataset
 barplot(table(cardiac$gender))
 
# Example2: Draw a histogram of the variable 'cholmmol' (cholesterol in mmol) in diabetes-dataset
 hist(diabetes$cholmmol)
 
# Draw a barplot of the variable 'gender' in diabetes-dataset
 barplot(table(diabetes$gender))
 
# Draw a histogram of the variable 'age' in cardiac-dataset
hist(cardiac$age)

# Draw a histogram of the variable 'bhr' (= "basal heart rate") in cardiac-dataset. Is there anything weird?
hist(cardiac$bhr)
 
# Draw a boxplot for cholesterol ('chol') in diabetes- data by gender.
boxplot(diabetes$chol~diabetes$gender)
 
# Draw a scatter-plot for the variables maximum heart rate ('maxhr') and systolic blood pressure ('sbp') in cardiac-data 
plot(cardiac$maxhr,cardiac$sbp)

# Check the correlation of maximum heart rate ('maxhr') and systolic blood pressure ('sbp') in cardiac-data.
cor(cardiac$maxhr, cardiac$sbp)
-0.04567112
```
 
*** =sct
```{r}
test_student_typed("barplot(table(diabetes$gender))", not_typed_msg = "Make sure that you have added the table-command inside the barplot-command.")
test_student_typed("hist(cardiac$age)", not_typed_msg = "Make sure that you have not made any spelling mistakes.")
test_student_typed("hist(cardiac$bhr)", not_typed_msg = "Make sure that you have not made any spelling mistakes.")
test_student_typed("boxplot(diabetes$chol~diabetes$gender)", not_typed_msg = "Make sure that you have not made any spelling mistakes.")
test_student_typed("plot(cardiac$maxhr,cardiac$sbp)", not_typed_msg = "Make sure that you have not made any spelling mistakes.")
test_output_contains("-0.04567112", incorrect_msg = "Make sure that you have checked the correlation coefficient.")

success_msg("Awesome! You old plotting wizard!")
```
--- type:NormalExercise lang:r xp:100 skills:1 key:exercise0_5

## Making plots more fancy
 
Plots are cooler with some colours in it. Title is a good way to tell people what the plot is about. One of R's best features is it wide range of colours, and easy wat of adjusting your plots to be exactly as you want. The basic colours, such as red and green can be gained by just typing 'red' or 'green'. To fin out more about colours in R, and their names, google colours in R and you are on your way to a super fancy plots!

*** =hint
Just add the title and colour options inside the plotting code, just like in the example.
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
*** =instructions
- Draw a barplot of the variable 'gender' in diabetes-dataset. Add a title and colour to you plot. Don't be afraid to get creative!
- Draw a histogram of the variable 'age' in cardiac-dataset. Add a title and colour to you plot.
- Draw a boxplot for cholesterol ('chol') in diabetes- data by gender. Add a title and different colours for the two boxes. 
 
*** =sample_code
```{r}
# Draw a barplot of the variable 'gender' in cardiac-dataset. Add a title and colour to you plot.
 
barplot(table(cardiac$gender),main="Gender in cardiac-dataset",col="tomato")

# Draw a barplot of the variable 'gender' in diabetes-dataset. Add a title and colour to you plot.
 

# Draw a histogram of the variable 'age' in cardiac-dataset. Add a title and colour to you plot.


# Draw a boxplot for cholesterol ('chol') in diabetes- data by gender. Add a title and different colours for the two boxes. Hint! use col=c("colour1","colour2")  
 
```
 
*** =solution
```{r}
# Draw a barplot of the variable 'gender' in diabetes-dataset
 
barplot(table(diabetes$gender),main="Title",col="red")
 
# Draw a barplot of the variable 'age' in cardiac-dataset
hist(cardiac$age,main="Title",col="red")

# Draw a boxplot for cholesterol ('chol') in diabetes- data by gender. Add a title and different colours for the two boxes. 
boxplot(diabetes$chol~diabetes$gender, col=c("red", "blue"))

```
 
*** =sct
```{r}
test_function("barplot", incorrect_msg = "Did you enter both title and colour options?")
test_function("hist", incorrect_msg = "Did you enter both title and colour options?")
success_msg("Oh your plots look just fabulous!")
```

