--- 
title_meta  : Chapter 0
title       : Sandbox
description : "In this chapter, you will learn the basics of how to work on DataCamp and how to handle with exercises. 
Since many of you are working with iPads, there will also be some advice regarding special characters such as `^` or `'`. Let's get started!"
 
--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_1
## THE WARM UP STARTS HERE: How does DataCamp work
 
In the box below you can see the instructions for the exercise. 

In the editor on the right you should type R code to solve the exercises. When you hit the 'Submit Answer' button, every line of code is interpreted and executed by R and you get a message whether or not your code was correct. The output of your R code is shown in the console in the lower right corner. You can also execute the code by pressing Ctrl+Enter (Cmd+Enter in Mac)

On the right in the upper window you can type your code. That window is called editor. 
You can execute the code by pressing Ctrl+Enter (Cmd+Enter in Mac) when the cursor is on the same line with your command,
and when you are done with the exercises, you can press the Submit-button.
If your command is on multiple lines you must select all those rows and then Ctrl+Enter or Cmd+Enter.
When you run your commands, the results appear in the window under the editor. That window is called console.

Tips for iPad users:
In case you are using an iPad, you can just go ahead and run all the commands by pressing the Submit-button. 
If you do this when you are not yet 
finished and some sort of error message appears, you can try to comment out the parts that you have not completed yet. 
That can be done by inserting a hashtag `#` at the beginning of the not completed lines.

*** =instructions

- In the editor on the right there is already some sample code. Can you see which lines are actual R code and which are comments?
- Add a line of code that calculates the sum of 4 and 18, and hit the 'Submit Answer' button.

*** =hint
Just add a line of R code that calculates the sum of 4 and 18, just like the example in the sample code!

*** =pre_exercise_code
```{r}
 # no pec
```

*** =sample_code
```{r}
# Calculate 3 + 4
3 + 4
# Calculate 4 + 18
```

*** =solution
```{r}
# Calculate 3 + 4
3 + 4
# Calculate 4 + 18
4 + 18
```

*** =sct
```{r}
test_output_contains("22", incorrect_msg = "Make sure to add `4 + 18` on a new line. Do not start the line with a `#`, otherwise your R code is not executed!")
success_msg("Awesome! See how the console shows the result of the R code you submitted? Now that you're familiar with the interface, let's get down to R business!")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:45129664b8
## Arithmetic with R

In its most basic form, R can be used as a simple calculator. Consider the following arithmetic operators:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^` or `**`

The last one might need some explaining:
- The `^` or `**` operator raises the number to its left to the power of the number to its right: for example `3^2` or `3**2` is 9.

With this knowledge, follow the instructions below to complete the exercise.

*** =instructions
- Type `2^5` in the editor after the "Exponentiation" part to calculate 2 to the power 5.
- Click 'Submit Answer' and have a look at the R output in the console.
- Note how the `#` symbol is used to add comments on the R code.

*** =hint
Make sure you have used either `^` or `**` signs.

*** =pre_exercise_code
```{r}
 # no pec
```

*** =sample_code
```{r}
# An addition
5 + 5 
# A subtraction
5 - 5 
# A multiplication
3 * 5
 # A division
(5 + 5) / 2 
# Exponentiation

```
