--- 
title_meta  : Chapter 3
title       : Statistical testing
description : "In this chapter, you will learn what does statistical significance mean and revise crosstabulation and learn to run some useful statistical tests, such as chi square and t-test."
 
--- type:NormalExercise lang:r xp:0 skills:1 key:ex2point1
## Hands on to t-test
 
Although in these days we have many tools for calculation and programming, it is good to understand the theory that lies behind R functions for example. t-test is quite a simple procedure and now we are going to learn how to run it without using the function that R offers. You just have to calculate the difference between two means and then divide it with the square root of (variance1 / n1) + (variance2 / n2)
 
*** =instructions
 - Find out whether there is a difference between men and women in basal heart rate by using t-test
    - Imagine we have a sample of 10 (n=10) people, 4 of which are women and 6 men  
    - Mean of women's basal heart rate is 78.5 and standard deviation is 6.65
    - Mean of men's basal heart rate is 75.67 and standard deviation is 20.93
    - The chosen significance level is 0.05
 - You can find the t-distribution table from the internet, just type 't distribution table' on google.
 - You will need to check the degrees of freedom (df) in order to use the distribution table (n1 + n2 - 2) 
 
*** =hint
Remember to use all necessary brackets. Square root in R is sqrt()
 
*** =pre_exercise_code
```{r}
 # no pec
```
 
*** =sample_code
```{r}
# Assign the number of women to n1 and number of men to n2
n1 <-
n2 <-
 
# Assign the mean value of basal heart rate of women to mean1 and standard deviation to sd1
mean1 <-
sd1 <-
 
# Do similarly to men (so assign values to mean2 and sd2)
 
 
# Now remember how to calculate the t-value? 
# If not, you can revise it from the course material or from the internet.
# Remember that variance is squared standard deviation.
# Now just calculate it here below and print the value
t <- 
t

# Calculate the degrees of freedom and print the value
df <- 
df

# Search for the right value from the distribution table
# The value is: 

# Does this support the alternative hypothesis (on significance level 0.05)? Remove the wrong answer.
# Your answer: YES / NO
```
 
*** =solution
```{r}
# Assign the number of women to n1 and number of men to n2
n1 <- 4
n2 <- 6
 
# Assign the mean value of basal heart rate of women to mean1 and standard deviation to sd1
mean1 <- 78.5
sd1 <- 6.65
 
# Do similarly to men
mean2 <- 75.67
sd2 <- 20.93
 
# Now remember how to calculate the t-value? 
# If not, you can revise it from the course material or from the internet.
# Remember that variance is squared standard deviation.
# Now just calculate it here below
t <- (mean1 - mean2) / (sqrt(((sd1^2) / n1) + ((sd2^2) / n2)))
t

# Calculate the degrees of freedom
df <- n1 + n2 - 2
df

# Search for the right value from the distribution table
# The value is: 1.860

# Does this support the alternative hypothesis (on significance level 0.05)? Remove the wrong answer.
# Your answer: NO
```
 
*** =sct
```{r}
test_student_typed("n1 <- 4", not_typed_msg = "Make sure you entered the right values given in the exercise.")
test_student_typed("mean1 <- 78.5", not_typed_msg = "Make sure you entered the right values given in the exercise.")
test_student_typed("sd1 <- 6.65", not_typed_msg = "Make sure you entered the right values given in the exercise.")
test_student_typed("mean2 <- 75.67", not_typed_msg = "Make sure you entered the right values given in the exercise.")
test_student_typed("sd2 <- 20.93", not_typed_msg = "Make sure you entered the right values given in the exercise.")
test_output_contains("0.3086563", incorrect_msg = "Make sure that you have studied well the formula for the t-value.")
test_output_contains("8", incorrect_msg = "Are you sure you calculated the degrees of freedom correctly?")
test_student_typed("1.86", not_typed_msg = "Are you sure you looked at the right spot on the distribution table?")
test_student_typed("NO", not_typed_msg = "Perhaps you need to check again how to interpret your results.")
success_msg("t-test seems to be your cup of tea! You know the name of the test actually originates from the lady who was convinced she can taste whether the milk or the tea was poured first on her teacup?")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:ex2point2
## Getting more familiar with the t-test 1
 
The t-test is often used to check whether there is differences between two independent groups, quite often men and women, sick and healthy or some other dicotomous groups. T-test can be easily done in R by using the function t.test(variable1~variable2)
 
*** =instructions
 - Run a t-test to compare systolic blood pressure between genders in cardiac-dataset. Is there any difference between men and women?

*** =hint
Use the R function t.test and just enter the variables.
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# The dataset cardiac is available
 
# Compare the systolic blood pressures by using t-test

# Is there a statistically significant difference? Remove the wrong answer
# Your answer: NO YES
 

```
 
*** =solution
```{r}
# The dataset cardiac is available
 
# Compare the systolic blood pressures by using t-test
with(cardiac,t.test(sbp~gender))

# Is there a statistically significant difference? Remove the wrong answer
# Your answer: NO

```
 
*** =sct
```{r}
test_student_typed("t.test", not_typed_msg = "Are you using R's function for t-test?")
test_student_typed("sbp~", not_typed_msg = "Make sure you have the variables inside the t.test-function in the right order.")
test_student_typed("NO", not_typed_msg = "Your interpretation of the t-test needs re-checking. Check whether the confidence interval contains zero or not.")
success_msg("t-test appears to be your cup of tea!")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:ex2point3
## Visualizing the results of the t-test
 
When you are doing statistical testing, it can be useful to visualize the results. In previous chapter you have learned many ways of plotting in R. To visualize the difference (or no difference) in means between two groups, boxplot is often a good choice. Now lets try to revise how it's done!
 
*** =instructions
 - Draw a boxplot to describe systolic blood pressure in males and females in cardiac-dataset. What can you see?

*** =hint
Use the R function boxplot and just enter the variables.
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# The dataset diabetes is available
# The dataset cardiac is available
 
# Visualize BMI in males and females in diabetes-dataset
with(diabetes, boxplot(bmi~gender, col="turquoise", main="BMI in males and females"))
 
# Visualize the systolic blood pressures in males and females in cardiac-dataset

axis(side = 1, at=1:2, labels=c("male","female")) # this adds text labels instead of 0 and 1. If you use this, then include command xaxt="n" in your boxplot command!
 
```
 
*** =solution
```{r}
# The dataset diabetes is available
# The dataset cardiac is available
 
# Visualize BMI in males and females in diabetes-dataset
with(diabetes, boxplot(bmi~gender, col="turquoise", main="BMI in males and females"))
 
# Visualize the systolic blood pressures in males and females in cardiac-dataset
with(cardiac, boxplot(sbp~gender, col="yellow", main="Systolic blood pressure in males and females", xaxt="n"))
axis(side = 1, at=1:2, labels=c("male","female")) # this adds text labels instead of 0 and 1. If you use this, then include command xaxt="n" in your boxplot command!
```
 
*** =sct
```{r}
test_student_typed("boxplot", not_typed_msg = "Are you using R's function for boxplot?")
success_msg("That's some nice looking boxplot! See how it highlights that there is no difference between the groups?")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:ex2point4
## Getting more familiar with the t-test 2
 
The t-test is often used to check whether there is differences between two independent groups, quite often men and women, sick and healthy or some other binary (two classes) groups.  T-test can be easily done in R by using the function t.test(variable1~variable2)
 
*** =instructions
 - Run a t-test to compare BMI between genders in diabetes-dataset. Is there any difference between men and women?

*** =hint
Use the R function t.test and just enter the variables.
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# The dataset diabetes is available
 
# Compare the BMIs between genders by using t-test

# Is there a statistically significant difference? Remove the wrong answer.
# Your answer: NO YES

```
 
*** =solution
```{r}
# The dataset diabetes is available
 
# Compare the BMIs between genders by using t-test
with(diabetes,t.test(bmi~gender))

# There is a statistically significant difference, do you agree? Remove the wrong answer.
# Your answer: YES

```
 
*** =sct
```{r}
test_student_typed("t.test", not_typed_msg = "Are you using R's function for t-test?")
test_student_typed("bmi~", not_typed_msg = "Make sure you have the variables inside the t.test-function in the right order.")
test_student_typed("YES", not_typed_msg = "Your interpretation of the t-test needs re-checking. Check whether the confidence interval contains zero or not.")
success_msg("Wow, you can already analyze the results of the t-test so well, congrats! Now let's move on to a next letter.")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:ex2point5
## The Chi square test 1
 
When dealing with discrete variables, especially binary ones (the ones with only two possible values), the test often chosen to check for independence between them, is the chi square test. The chi square test in R can be done simply with the function chisq.test(table(variable1,variable2))

*** =instructions
 - Crosstabulate event and restwma from the cardiac-dataset.
   - Run the chi square test to check whether those two variables depend on each other.
   
*** =hint
First make a table and then put the table inside the chi square function. The chi square test in R is chisq.test()
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# The dataset cardiac is available
# HINT! In order to see the variable names in table output, you can use command with(nameofyourdata, table(variable1,variable2)) as instructed in previous chapters
# HINT! You can use the with-function in all other codes too, such as with(nameofyourdata, plot(variable1,variable2)

# Crosstabulate event and restwma
table1 <-

# Run the chi square test for the table you just created
test1 <- 
test1

# Print out the p-value of the test separately
test1$p.value

# Do the variables depend on each other? Remove the wrong answer.
# Your answer: NO YES

```
 
*** =solution
```{r}
# The datasets cardiac and diabetes are available
# HINT! In order to see the variable names in table output, you can use command with(nameofyourdata, table(variable1,variable2))
# HINT! You can use the with-function in all other codes too, such as with(nameofyourdata, plot(variable1,variable2)

# Crosstabulate event and restwma
table1 <- with(cardiac, table(event, restwma))

# Run the chi square test for the table you just created
test1 <- chisq.test(table1)

# Print out the p-value of the test separately
test1$p.value

# Do the variables depend on each other? Remove the wrong answer.
# Your answer: YES
```
 
*** =sct
```{r}
test_student_typed("event", not_typed_msg = "Make sure you entered the right variables in your command.")
test_student_typed("restwma", not_typed_msg = "Make sure you entered the right variables in your command.")
test_student_typed("YES", not_typed_msg = "Make sure your interpretation of the first chi square test result is correct.")
test_function("chisq.test", incorrect_msg = "Are you using the right test function?")
test_output_contains("3.359", incorrect_msg = "Your chi square value seems to be incorrect. Look at your code carefully, maybe you used a wrong variable?")
success_msg("That's some badass crosstabulation you did there, go you!")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:ex2point6
## Visualizing the results of the Chi square
 
As we discussed earlier, when you are doing statistical testing, it can be useful to visualize the results. In previous chapter you have learned many ways of plotting in R. To visualize the dependence between two discrete variables it could be useful to make a stacked barplot, where you have bars representing groups in the other variables and the bar heights representing the percentage of the gorups in the other variable. Take a look at the example code -->
 
*** =instructions
 - Draw a stacked barplot (scaled to 100 %) of variables event and restwma in cardiac-dataset. What can you see?
  - First transform your table into % as in the example code. 
  - Then draw a barplot.
  - Add legend to your plot.
   - If you want to change your legend somehow, take a look at the help by typing ?legend to see the options.

*** =hint
Take a look at the example code and just enter your variables.
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# The dataset diabetes is available
# The dataset cardiac is available

# Draw a stacked barplot (scaled to 100 %) of variables smoking status and death in cardiac-dataset

# Transform this data in %
data_percentage=apply(table(cardiac$death,cardiac$smoking), 
                      2, function(x){x*100/sum(x,na.rm=T)})

# Make a stacked barplot--> notice it will be in %!
barplot(data_percentage, col=c("turquoise","cornflowerblue") , 
        xlab="smoking status\n (0 = non-smoker, 1 = smoker)", ylab="%", main="Percentages of deaths in smoking groups\n\n")
legend("bottomright", c("no death", "death"), 
       fill=c("turquoise","cornflowerblue"),
       inset=c(0,1), xpd=TRUE,  horiz=TRUE, bty="n")

# Draw a stacked barplot (scaled to 100 %) of variables event and restwma in cardiac-dataset

# Transform this data in %


# Make a stacked barplot--> notice it will be in %!

 
```
 
*** =solution
```{r}
# Draw a stacked barplot (scaled to 100 %) of variables death and anyevent in cardiac-dataset

# Draw a stacked barplot (scaled to 100 %) of variables smoking status and death in cardiac-dataset

# Transform this data in %
data_percentage=apply(table(cardiac$death,cardiac$smoking), 
                      2, function(x){x*100/sum(x,na.rm=T)})

# Make a stacked barplot--> it will be in %!
barplot(data_percentage, col=c("turquoise","cornflowerblue") , 
        xlab="smoking status\n (0 = non-smoker, 1 = smoker)", ylab="%", main="Percentages of deaths in smoking groups\n\n")
legend("bottomright", c("no death", "death"), 
       fill=c("turquoise","cornflowerblue"),
       inset=c(0,1), xpd=TRUE,  horiz=TRUE, bty="n")

# Draw a stacked barplot (scaled to 100 %) of variables event and restwma in cardiac-dataset

# Transform this data in %
data_percentage=apply(table(cardiac$event,cardiac$restwma), 
                      2, function(x){x*100/sum(x,na.rm=T)})

# Make a stacked barplot--> it will be in %!
barplot(data_percentage, col=c("turquoise","cornflowerblue") , 
        xlab="restwma\n (0 = no anomalies, 1 = anomalies)", ylab="%", main="Percentages of events in restwma-groups\n\n")
legend("bottomright", c("no event", "event"), 
       fill=c("turquoise","cornflowerblue"),
       inset=c(0,1), xpd=TRUE,  horiz=TRUE, bty="n")
```
 
*** =sct
```{r}
test_student_typed("barplot(", not_typed_msg = "Are you using R's function for barplot?")
test_student_typed("legend", not_typed_msg = "Did you remember to insert the legend?")
success_msg("Nice colors! Does there seem to be difference between the groups? Does that match the result of your chi square test")
```


--- type:NormalExercise lang:r xp:0 skills:1 key:ex2point7
## The Chi square test 2
 
When dealing with discrete variables, especially dicotomous ones, the test often chosen to check for independence is the chi square test.  The chi square test in R can be done simply with the function chisq.test(table(variable1,variable2))

*** =instructions
 - Crosstabulate smoking and gender from the cardiac-dataset. 
   - Run the chi square test to check whether those two variables depend on each other.
  
*** =hint
First make a table and then put the table inside the chi square function. The chi square test in R is chisq.test()
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# The dataset cardiac is available.

# Crosstabulate smoking and gender
table2 <-

# Run the chi square test for the table you just created
test2 <-
test2

# Print out the p-value of the test separately
test2$p.value

# Do the variables depend on each other? Remove the wrong answer.
# Your answer: NO YES

```
 
*** =solution
```{r}
# The dataset cardiac is available.

# Crosstabulate smoking and gender
table2 <- with(cardiac, table(smoking, gender))

# Run the chi square test for the table you just created
test2 <- chisq.test(table2)

# Print out the p-value of the test separately
test2$p.value

# Do the variables depend on each other? Remove the wrong answer.
# Your answer: YES
```
 
*** =sct
```{r}
test_student_typed("smoking", not_typed_msg = "Make sure you entered the right variables in your command.")
test_student_typed("gender", not_typed_msg = "Make sure you entered the right variables in your command.")
test_student_typed("YES", not_typed_msg = "Make sure your interpretation of the second chi square test result is correct.")
test_function("chisq.test", incorrect_msg = "Are you using the right test function?")
test_output_contains("4.79", incorrect_msg = "Your chi square value seems to be incorrect. Look at your code carefully, maybe you used a wrong variable?")
success_msg("Way to go, you are getting more and more pro at this!")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:ex2point8
## The Chi square test 3
 
When dealing with discrete variables, especially dicotomous ones, the test often chosen to check for independence is the chi square test. The chi square test in R can be done simply with the function chisq.test(table(variable1,variable2))

*** =instructions
 - Crosstabulate categorized cholesterol (cholcat) and gender from the diabetes-dataset. 
   - Run the chi square test to check whether those two variables depend on each other.
   
*** =hint
First make a table and then put the table inside the chi square function. The chi square test in R is chisq.test()
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# The dataset diabetes is available

# Crosstabulate categorized cholesterol and gender
table3 <-

# Run the chi square test for the table you just created
test3 <-
test3

# Print out the p-value of the test separately
test3$p.value

# Do the variables depend on each other? Remove the wrong answer.
# Your answer: NO YES
 
```
 
*** =solution
```{r}
# The datasets diabetes is available

# Crosstabulate categorized cholesterol and gender
table3 <- with(diabetes, table(cholcat, gender))

# Run the chi square test for the table you just created
test3 <- chisq.test(table3)

# Print out the p-value of the test separately
test3$p.value

# Do the variables depend on each other? Remove the wrong answer.
# Your answer: NO 
 
```
 
*** =sct
```{r}
test_student_typed("gender", not_typed_msg = "Make sure you entered the right variables in your command.")
test_student_typed("cholcat", not_typed_msg = "Make sure you entered the right variables in your command.")
test_student_typed("NO", not_typed_msg = "Make sure your interpretation of the third chi square test result is correct.")
test_function("chisq.test", incorrect_msg = "Are you using the right test function?")
test_output_contains("0.85", incorrect_msg = "Your third chi square value seems to be incorrect. Look at your code carefully, maybe you used a wrong variable?")
success_msg("Be there or be chi square(d)! You definately are there, well done with your chi square tests!")
```

--- type:MultipleChoiceExercise lang:r xp:0 skills:1 key:ex2point9
## Remembering what was the t-test all about
 
What is Student's t-test used for?
 
*** =instructions
- To check for missingness across all variables in the data.
- Counting the variables that are discrete.
- Checking whether the mean of a variable is equal in two populations.
- Calculating the median of a variable.

*** =hint
Its name might give you a hint that t-test is a test. So we are testing whether something is true or not.
 
*** =pre_exercise_code
```{r}
 # no pec
```
 
*** =sample_code
```{r}

```
 
*** =solution
Checking whether the mean of a variable is equal in two populations

 
*** =sct
msg1 <- "Not quite there yet!"
msg2 <- "There is no shame in googling, bro."
msg3 <- "That's right, well done!"
msg4 <- "Come on, you can do better!"
test_mc(3, feedback_msgs = c(msg1, msg2, msg3, msg4))

--- type:MultipleChoiceExercise lang:r xp:0 skills:1 key:ex2point10
## True or not about the t-test?
 
Which of the following is true?
 
*** =instructions
- The t-statistic was invented by William van der Gosset who worked at Heineken brewery at the beginning of 1900s. 
- The independent samples t-test could be done for example if one wanted to compare the heights of a certain group of school children at age 6 and then again at age 10.
- The t-test is especially used when we have only binary variables in our data.
- The t-test could be used to compare the BMI between the people born in winter and people born in summer. 

*** =hint
Think about what you have done with the t-test in the previous exercises.
 
*** =pre_exercise_code
```{r}
 # no pec
```
 
*** =sample_code
```{r}

```
 
*** =solution
The t-test could be used to compare the BMI between the people born in winter and people born in summer.

 
*** =sct
msg1 <- "It was indeed invented by a certain William that worked in a brewery, but that wasn't Heineken. Google if interested ;)"
msg2 <- "If we are measuring the same group of people again, then we are not talking about independent, but dependent samples."
msg3 <- "The t-test is often used to compare means, and means shouldn't be even calculated for binary variables."
msg4 <- "Yes, that is correct! The t-test can be used to compare for example the BMI between two independent groups."
test_mc(4, feedback_msgs = c(msg1, msg2, msg3, msg4))

--- type:MultipleChoiceExercise lang:r xp:0 skills:1 key:ex2point11
## Do you still remember the Chi square?
 
What is the Chi square test usually used for?
 
*** =instructions
- To test for independence between variables. 
- To test whether the data is continuous or not.
- To find the upper and lower quartile of both variables.
- To check if a certain observation has a certain variable value. 

*** =hint
Now we are talking about the chi square test that is often used after crosstabulating variables.
 
*** =pre_exercise_code
```{r}
 # no pec
```
 
*** =sample_code
```{r}

```
 
*** =solution
To test for independence between variables.

 
*** =sct
msg1 <- "Awesome, you knew it!"
msg2 <- "Chi square usually doesn't test the whole data at once."
msg3 <- "Usually the chi square test isn't used to find summary statistics for a variable."
msg4 <- "Usually the chi square test isn't used to find check for single observations."
test_mc(1, feedback_msgs = c(msg1, msg2, msg3, msg4))


--- type:MultipleChoiceExercise lang:r xp:0 skills:1 key:ex2point12
## Interpret Chi square test results
 
Type print(mydata) on the console and interpret whether there is independence between variables. (significance level 0.05) Hint: Perform a chi square test, either by hand or in console.
 
*** =instructions
- The p-value of the chi square test is > 0.05, so the variables don't seem to depend on each other.
- The p-value of the chi quare test is > 0.05, so the variables seem to depend on each other.
- The p-value of the chi square test is < 0.05, so the variables seem to depend on each other. 
- The p-value of the chi square test is < 0.05, so the variables don't seem to depend on each other.

*** =hint
Think first about what are we testing. Then think about the p-value: Does it have to be lower or greater than 0.05 for the test to be significant?
 
*** =pre_exercise_code
```{r}
a<-c(15,38)
b<-c(97,120)
mydata<-cbind(a,b)
```
 
*** =sample_code
```{r}

```
 
*** =solution
The p-value of the chi square test is < 0.05, so the variables seem to depend on each other. 

 
*** =sct
msg1 <- "Is the p-value over 0.05 though?"
msg2 <- "Maybe revise the interpretation of the p-value."
msg3 <- "Very well interpreted: The low p-value indicates that the test result is statistically significant, and since the null hypotheses of the chi square test is that the variables don't depend on each other, we can now reject that hypothesis."
msg4 <- "What is the null hypotheses in the chi square test?"
test_mc(3, feedback_msgs = c(msg1, msg2, msg3, msg4))


