---
title_meta: 'Chapter 2'
title: Plotting
description: 'In this chapter, you will learn to plot in R.'
---

## Basics of plotting and barplot- command

```yaml
type: NormalExercise
key: exercise2_1
lang: r
xp: 0
skills: 1
```

One of R's best features is its plotting features. In R it is easy and fun to adjust your plots just the way you like it. But now, let's start from the basics. Propably one of the first things you want to do with your data set is to check your gender distibution. The best way to do that is to draw a barplot from your variable 'gender'. To draw a barplot, use barplot(table(data$variable)).  Why table you may ask? What barplot wants inside of it is the heights of the bars, so if you would type in barplot(data$variable), it would produce as many bars as you have individuals.

`@instructions`
- Draw a barplot of the variable 'gender' in diabetes-dataset.

- Draw a barplot of the variable 'gender' in cardiac-dataset.

`@hint`
Remember to put your variable inside table, and refer to your variable as data$variable.

`@pre_exercise_code`
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```

`@sample_code`
```{r}
# Draw a barplot of the variable 'gender' in diabetes-dataset.

# Draw a barplot of the variable 'gender' in cardiac-dataset. 

```

`@solution`
```{r}
# Draw a barplot of the variable 'gender' in diabetes-dataset.
barplot(table(diabetes$gender))
# Draw a barplot of the variable 'gender' in cardiac-dataset. 
barplot(table(cardiac$gender))
```

`@sct`
```{r}
test_student_typed("barplot(table(diabetes$gender))", not_typed_msg = "Remember to put your variable inside table, and refer to your variable as data$variable. Also make sure you referred to right data set.")
test_student_typed("barplot(table(cardiac$gender))", not_typed_msg = "Remember to put your variable inside table, and refer to your variable as data$variable. Also make sure you referred to right data set.")
success_msg("Hey, got some fancy looking plots over there! Now Let's make them even fancier.")
```

---

## More plotting commands

```yaml
type: NormalExercise
key: exercise2_2
lang: r
xp: 0
skills: 1
```

In previous exercise you learned to do a barplot from discrete variable. There is however many more other possibilites, such as histograms, pie charts, scatter plots, box plots etc. Now you will learn to do couple of them. Here is a list of a few most commonly used commands:
- histogram: hist(data$variable)
- scatter plot: plot(data$variable1, data$variable1)
- pie chart: pie(table(data$variable)) Note! pie- command works as barplot-command in previous exercise
- box plot: boxplot(data$variable)

`@instructions`
- Draw histogram from the variable 'age' in diabetes- data set.

- Draw scatter plot of variables 'sbp' (systolic blood pressure) and 'maxhr' (maximum heart rate) in cardiac data set.

- Draw pie chart from variable 'gender' in cardiac- data set

- Draw a boxplot from variable 'age' in diabetes- data set 

`@hint`
Use commands hist(), plot(), pie(), and boxplot(). Make sure to put right variables in right format.

`@pre_exercise_code`
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```

`@sample_code`
```{r}
# Draw histogram from the variable 'age' in diabetes data set:

# Draw scatter plot of variables 'sbp' and 'maxhr' in cardiac data set:

# Draw pie chart from variable 'gender' in cardiac- data set:

# Draw a boxplot from variable 'age' in diabetes- data set:

```

`@solution`
```{r}
# Draw histogram from the variable 'age' in diabetes data set
hist(diabetes$age)

# Draw scatter plot of variables 'sbp' and 'maxhr' in cardiac data set. 
plot(cardiac$sbp, cardiac$maxhr)

# Draw pie chart from variable 'gender' in cardiac- data set:
pie(table(cardiac$gender))

# Draw a boxplot from variable 'age' in diabetes- data set:
boxplot(diabetes$age)

```

`@sct`
```{r}
test_student_typed("hist(diabetes$age)", not_typed_msg = "Make sure you used hist()- command. Also make sure you referred to right data set and variable.")
test_student_typed("plot(", not_typed_msg = "Make sure you used plot()- command. Also make sure you referred to right data set and variables.")
test_student_typed("pie(table(cardiac$gender))", not_typed_msg = "Make sure you used pie()- command, and put your variable inside table()-command. Also make sure you referred to right data set and variable.")
test_student_typed("boxplot(diabetes$age)", not_typed_msg = "Make sure you used boxlot()- command. Also make sure you referred to right data set and variable.")
success_msg("Nice one! You are about to become a real plotting wizard!")
```


---
## Adding some colour!

```yaml
type: NormalExercise
key: exercise2_4
lang: r
xp: 0
skills: 1
```

R has very verstatile selection of colours. Basic colours, such as red and green can be obtained by typing "red" or "green". To learn more  colours: check http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf. To add colour into your plot, add argument col = "" inside your plotting command. If you want to add multiple colours, for example pie chart for gender, just give col as many colours you want inside vector (c()). For example: col = c("red", "green").

`@instructions`
- Draw scatter plot of variables 'sbp' (systolic blood pressure) and 'maxhr' (maximum heart rate) in cardiac data set, and make your dots yellow.

- Draw pie chart from variable 'gender' in cardiac- data set, and make your sections red and blue.



`@hint`
Use command plot() and pie ()to create your plot. Add correct colours to your plots
`@pre_exercise_code`
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```

`@sample_code`
```{r}
# - Draw scatter plot of variables 'sbp' (systolic blood pressure) and 'maxhr' (maximum heart rate) in cardiac data set, and make your dots yellow.

- Draw pie chart from variable 'gender' in cardiac- data set, and make your sections red and blue.


```

`@solution`
```{r}
# - Draw scatter plot of variables 'sbp' (systolic blood pressure) and 'maxhr' (maximum heart rate) in cardiac data set, and make your dots yellow.

plot(cardiac$sbp, cardiac$maxhr, col = "yellow")

- Draw pie chart from variable 'gender' in cardiac- data set, and make your sections red and blue.
pie(table(cardiac$gender), col = c("red", "blue"))
```

`@sct`
```{r}
test_student_typed("pie(table(cardiac$gender)", not_typed_msg = "Make sure you used pie()- command. Also make sure you referred to right data set and variable.")
success_msg("Nice one! You are about to become a real plotting wizard!")
```
