--- 
title_meta  : Chapter 0
title       : Sandbox
description : "In this chapter, you will learn the basics of how to work on DataCamp and how to handle with exercises. 
Since many of you are working with iPads, there will also be some advice regarding special characters such as `^` or `'`."
 
--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_1
## THE WARM UP STARTS HERE: How does DataCamp work
 
In the box below you can see the instructions for the exercise. 
On the right in the upper window you can type your code. That window is called editor. 
You can execute the code by pressing Ctrl+Enter (Cmd+Enter in Mac) when the cursor is on the same line with your command,
and when you are done with the exercises, you can press the Submit-button.
If your command is on multiple lines you must select all those rows and then Ctrl+Enter or Cmd+Enter.
When you run your commands, the results appear in the window under the editor. That window is called console.

Tips for iPad users:
In case you are using an iPad, you can just go ahead and run all the commands by pressing the Submit-button. 
If you do this when you are not yet 
finished and some sort of error message appears, you can try to comment out the parts that you have not completed yet. 
That can be done by inserting a hashtag `#` at the beginning of the not completed lines.

--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_2
## Arithmetic with R
 
You can use R as a calculator. You can run basic mathematical operations like this: 
 
- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`

With this knowledge, follow the instructions below to complete the exercise.

`@instructions`
- Type `2^5` in the editor to calculate 2 to the power 5.
- Click 'Submit Answer' and have a look at the R output in the console.
- Note how the `#` symbol is used to add comments on the R code.

`@hint`
Another example of the modulo operator: `9 %% 2` equals `1`.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# An addition
5 + 5 
# A subtraction
5 - 5 
# A multiplication
3 * 5
 # A division
(5 + 5) / 2 
# Exponentiation
# Modulo
```

`@solution`
```{r}
# An addition
5 + 5
# A subtraction
5 - 5 
# A multiplication
3 * 5
 # A division
(5 + 5) / 2 
# Exponentiation
2 ^ 5
# Modulo
28 %% 6
```

`@sct`
```{r}
msg = "Do not remove the other arithmetic examples!"
test_output_contains("2^5", incorrect_msg = "The exponentiation example is not correct. Write `2 ^ 5` on a new line.")
test_output_contains("28 %% 6", incorrect_msg = "There seems to be an issue with the modulo example. Write `28 %% 6` on a new line.")
success_msg("Great! Head over to the next exercise.")
```
