--- 
title_meta  : Chapter 0
title       : Sandbox
description : "In this chapter, you will learn the basics of how to work on DataCamp and how to handle with exercises. 
Since many of you are working with iPads, there will also be some advice regarding special characters such as `^` or `'`."
 
--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_1
## THE WARM UP STARTS HERE: How does DataCamp work
 
In the box below you can see the instructions for the exercise. 
On the right in the upper window you can type your code. That window is called editor. 
You can execute the code by pressing Ctrl+Enter (Cmd+Enter in Mac) when the cursor is on the same line with your command,
and when you are done with the exercises, you can press the Submit-button.
If your command is on multiple lines you must select all those rows and then Ctrl+Enter or Cmd+Enter.
When you run your commands, the results appear in the window under the editor. That window is called console.

Tips for iPad users:
In case you are using an iPad, you can just go ahead and run all the commands by pressing the Submit-button. 
If you do this when you are not yet 
finished and some sort of error message appears, you can try to comment out the parts that you have not completed yet. 
That can be done by inserting a hashtag `#` at the beginning of the not completed lines.

--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_2
## Arithmetic with R
 
You can use R as a calculator. You can run basic mathematical operations like this: 
 
- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`

With this knowledge, follow the instructions below to complete the exercise.

`@instructions`
- Type `2^5` in the editor to calculate 2 to the power 5.
- Click 'Submit Answer' and have a look at the R output in the console.
- Note how the `#` symbol is used to add comments on the R code.

`@hint`
Another example of the modulo operator: `9 %% 2` equals `1`.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# An addition
5 + 5 
# A subtraction
5 - 5 
# A multiplication
3 * 5
 # A division
(5 + 5) / 2 
# Exponentiation
# Modulo
```

`@solution`
```{r}
# An addition
5 + 5
# A subtraction
5 - 5 
# A multiplication
3 * 5
 # A division
(5 + 5) / 2 
# Exponentiation
2 ^ 5
# Modulo
28 %% 6
```

`@sct`
```{r}
msg = "Do not remove the other arithmetic examples!"
test_output_contains("2^5", incorrect_msg = "The exponentiation example is not correct. Write `2 ^ 5` on a new line.")
test_output_contains("28 %% 6", incorrect_msg = "There seems to be an issue with the modulo example. Write `28 %% 6` on a new line.")
success_msg("Great! Head over to the next exercise.")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_3
## Save yout work
 
In order to understand better your own code, or in case you want to save it as an actual R script 
(It is possible as a simple copy-paste if you have R installed in your computer.) it is good to comment on your code. 
At the beginning when the code is still not very clear to you, it helps you to understand what the code does. When the time
goes by, these weird codewords start to mean something to you and it becomes easier, but the comments still help you
and are especially useful if you need to share your code with another person.
Commenting can be done by typing `#` at the beginning of the row. Any text you type on that row will be considered as a comment. 
For example _Write your name here_ in the editor window is a comment.

Don't be afraid of making mistakes, you do not loose any points by submitting an incorrect solution. 
If your solution is not correct, you will receive an error message and you can simply try again.
If, on the other hand you decide to take a hint, you will loose some points and if you decide to look at the ready-made 
solution, you will loose your points for that exercise.
 
*** =instructions
 - These first exercises are a warm up to get to know how does this environment work.
 - Whenever the commented line asks you to do something, you should write the answer under that line unless told otherwise.

*** =hint
Just follow the instructions written in the editor. If you do exactly as they say, you should be able to complete this
exercise. Try submitting your code and maybe the error messages help you to continue. If this doesn't help, ask the course
staff or a friend or try posting your question on Moodle, maybe there are other students struggling with the same problem!
 
*** =pre_exercise_code
```{r}
 # no pec
```
 
*** =sample_code
```{r}
# Write your name here

 
# Write here numbers from one to ten without spaces

 
# Calculate the sum of the following numbers: 1,3,5,7,9

 
# If Matti has 100 apples and he gives Mari 35 of them, how much apples does Matti have left? Calculate.
 

```
 
*** =solution
```{r}
# Write your name here
Maija Meikalainen
 
# Write here numbers from one to ten without spaces
12345678910
 
# Calculate the sum of the following numbers: 1,3,5,7,9
1+3+5+7+9
 
# If Matti has 100 apples and he gives Mari 35 of them, how much apples does Matti have left? Calculate.
100-35
```
 
*** =sct
```{r}
test_output_contains("12345678910", incorrect_msg = "When you are writing the numbers from one to ten, make sure there are no spaces between them.")
test_output_contains("25", incorrect_msg = "Your sum is not quite correct, try again.")
test_output_contains("65", incorrect_msg = "Try counting the apples again.")
success_msg("It seems that you are getting the idea. Let's still continue the warm up for a little bit.")
```
