--- 
title_meta  : Chapter 3.1
title       : Linear regression
description : "This weeks exercises will focus on linear regression. You will form a few models yourself and train to interpret the results from those models. You will also learn to evaluate the assumptions of the models."
 
--- type:NormalExercise lang:r xp:100 skills:1 key:3d083aaefb
## Linear regression with one explanatory variable
 
In this exercise you will form a linear regression model, see the results and make some interpretations based on those. In R linear regression can be done using lm()- function, for example: lm(cardiac$dobEF~cardiac$baseEF). Before the ~ sign you can put only one variable, the outcome variable. After it you can put one or multiple explanatory variables devided by the + sign. 
 
*** =instructions
 - Form a linear regression model, in which you will have systolic blood pressure (sbp) as an outcome variable and age as an explanatory variable. Use the Cardiac-data. 

 - Interpret the results: 
 - What is the regression coefficient?
 - What is the intercept?
 - What is the variance explained?
 
*** =hint
Use formula lm(data$outcome~data$explanatory). To show the results, use summary-fucntion. 
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# Form the model:

# Check the results:

# Fill the following sentences with the right number:
# Regression coefficient is:
# Intercept is:
# Variance explained is:

```
 
*** =solution
```{r}
reg1<-lm(cardiac$sbp~cardiac$age)
summary(reg1)
# Fill the following sentences with the right number:
# Regression coefficient is: -0.2203
# Intercept is: 161.7502
# Variance explained is: 0.00349
```
 
*** =sct
```{r}
test_output_contains("lm(cardiac$sbp~cardiac$age)", incorrect_msg = "Make sure you have put the variables in the right order, and they are divided by the ~ sign.")
test_student_typed("summary(", not_typed_msg = "Make sure you have used the summary-function to see the results.")
test_student_typed("Regression coefficient is: -0.22", not_typed_msg = "Make sure you have typed the right regression coefficient.")
test_student_typed("Intercept is: 161.7", not_typed_msg = "Make sure you have typed the right intercept.")
test_student_typed("Variance explained is: 0.00349", not_typed_msg = "Make sure you have typed the right variance explained.")


```

--- type:NormalExercise lang:r xp:100 skills:1 key:harj32
## Linear regression with multiple explanatory variables
 
In the previous exercise you made a linear regression using one explanatory variables. In this exercise you will add explanatory variables. 
 
*** =instructions
 - Form a linear regression model, in which you will have systolic blood pressure (sbp) as an outcome variable and age, sex and ejection fraction (baseEF) as an explanatory variable. Use the Cardiac-data. 
 Interpret the results: 
 - What is the regression coefficient?
 - What is the intercept?
 - What is the variance explained?
 
*** =hint
Use formula lm(data$outcome~data$explanatory1+data$explanatory2+ etc).  
 
*** =pre_exercise_code
```{r}
cardiac<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Cardiac.txt",header=T,row.names=1)
diabetes<-read.table("https://raw.githubusercontent.com/paulabergman/intro-to-biostatistics-with-r/master/Diabetes.txt",header=TRUE,row.names=1)
```
 
*** =sample_code
```{r}
# Form the model

# Check the results

# Regression coefficient for age

# Regression coefficient for sex

# Regression coefficient for ejection fraction

# Intercept for age

# Variance explained

```
 
*** =solution
```{r}
reg1<-lm(cardiac$sbp~cardiac$age+cardiac$gender+cardiac$baseEF)
summary(reg1)
# Regression coefficient for age
-0.2219
# Regression coefficient for gender
-0.1973
# Regression coefficient for ejection fraction
0.4314
# Intercept
137.9916
# Variance explained
0.001469
```
 
*** =sct
```{r}
test_output_contains("lm(cardiac$sbp~cardiac$age+cardiac$gender+cardiac$baseEF)", incorrect_msg = "Make sure you have typed all the variables separated by + sign, and their names correctly. ")
test_output_contains("summary("", incorrect_msg = "Make sure you have used the summary-function to see the results.")

```
