--- 
title_meta  : Chapter 0
title       : Sandbox
description : "In this chapter, you will learn the basics of how to work on DataCamp and how to handle with exercises. 
Since many of you are working with iPads, there will also be some advice regarding special characters such as `^` or `'`. Let's get started!"
 
--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_1
## THE WARM UP STARTS HERE: How does DataCamp work
 
In the box below you can see the instructions for the exercise. 

In the editor on the right you should type R code to solve the exercises. When you hit the 'Submit Answer' button, every line of code is interpreted and executed by R and you get a message whether or not your code was correct. The output of your R code is shown in the console in the lower right corner. You can also execute the code by pressing Ctrl+Enter (Cmd+Enter in Mac)

On the right in the upper window you can type your code. That window is called editor. 
You can execute the code by pressing Ctrl+Enter (Cmd+Enter in Mac) when the cursor is on the same line with your command,
and when you are done with the exercises, you can press the Submit-button.
If your command is on multiple lines you must select all those rows and then Ctrl+Enter or Cmd+Enter.
When you run your commands, the results appear in the window under the editor. That window is called console.

Tips for iPad users:
In case you are using an iPad, you can just go ahead and run all the commands by pressing the Submit-button. 
If you do this when you are not yet 
finished and some sort of error message appears, you can try to comment out the parts that you have not completed yet. 
That can be done by inserting a hashtag `#` at the beginning of the not completed lines.

*** =instructions

- In the editor on the right there is already some sample code. Can you see which lines are actual R code and which are comments?
- Add a line of code that calculates the sum of 4 and 18, and hit the 'Submit Answer' button.

*** =hint
Just add a line of R code that calculates the sum of 4 and 18, just like the example in the sample code!

*** =pre_exercise_code
```{r}
 # no pec
```

*** =sample_code
```{r}
# Calculate 3 + 4
3 + 4
# Calculate 4 + 18
```

*** =solution
```{r}
# Calculate 3 + 4
3 + 4
# Calculate 4 + 18
4 + 18
```

*** =sct
```{r}
test_output_contains("22", incorrect_msg = "Make sure to add `4 + 18` on a new line. Do not start the line with a `#`, otherwise your R code is not executed!")
success_msg("Awesome! See how the console shows the result of the R code you submitted? Now that you're familiar with the interface, let's get down to R business!")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_2
## Arithmetic with R

In its most basic form, R can be used as a simple calculator. Consider the following arithmetic operators:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^` or `**`

The last one might need some explaining:
- The `^` or `**` operator raises the number to its left to the power of the number to its right: for example `3^2` or `3**2` is 9.

With this knowledge, follow the instructions below to complete the exercise.

*** =instructions
- Type `2^5` in the editor after the "Exponentiation" part to calculate 2 to the power 5.
- Click 'Submit Answer' and have a look at the R output in the console.
- Note how the `#` symbol is used to add comments on the R code.

*** =hint
Make sure you have used either `^` or `**` signs.

*** =pre_exercise_code
```{r}
 # no pec
```

*** =sample_code
```{r}
# An addition
5 + 5 
# A subtraction
5 - 5 
# A multiplication
3 * 5
 # A division
(5 + 5) / 2 
# Exponentiation

```

*** =solution
```{r}
# An addition
5 + 5 
# A subtraction
5 - 5 
# A multiplication
3 * 5
 # A division
(5 + 5) / 2 
# Exponentiation
5^2
```

*** =sct
```{r}
msg = "Do not remove the other arithmetic examples!"
test_output_contains("25", incorrect_msg = "The exponentiation example is not correct. Write `5 ^ 2` or `5**2` on a new line.")
success_msg("Great! Head over to the next exercise.")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_3
## Variable assignment 

A basic concept in (statistical) programming is called a **variable**. 

A variable allows you to store a value (e.g. 4) or an object (e.g. a function description) in R. You can then later use this variable's name to easily access the value or the object that is stored within this variable. 

You can assign a value 3 to a variable `my_var` with the command

```
my_var <- 3
```

*** =instructions
Over to you: complete the code in the editor such that it assigns the value 42 to the variable `x` in the editor. Click 'Submit Answer'. Notice that when you ask R to print `x`, the value 42 appears.

*** =hint
Look at how the value 3 was assigned to `my_var` in the exercise's assignment. Do the exact same thing in the editor, but now assign 42 to the variable `x`.

*** =pre_exercise_code
```{r}
 # no pec
```

*** =sample_code
```{r}
# Assign the value 42 to x
x <- 
# Print out the value of the variable x
x
```

*** =solution
```{r}
# Assign the value 42 to x
x <- 42
# Print out the value of the variable x
x
```

*** =sct
test_object("x", undefined_msg = "Make sure to define a variable `x`.",
            incorrect_msg = "Make sure that you assign the correct value to `x`.") 
success_msg("Good job! Have you noticed that R does not print the value of a variable to the console when you did the assignment? `x <- 42` did not generate any output, because R assumes that you will be needing this variable in the future. Otherwise you wouldn't have stored the value in a variable in the first place, right? Proceed to the next exercise!")

--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_4
## Variable assignment (2)

Suppose you have a fruit basket with seven oranges. As a data analyst in training, you want to store the number of oranges in a variable with the name `my_oranges`. 

*** =instructions
- Type the following code in the editor: `my_oranges <- 7`. This will assign the value 7 to `my_oranges`.
- Type: `my_oranges` below the second comment. This will print out the value of `my_oranges`.
- Click 'Submit Answer', and look at the console: you see that the number 7 is printed. So R now links the variable `my_oranges` to the value 7.

*** =hint
Remember that if you want to assign a number or an object to a variable in R, you can make use of the assignment operator `<-`. Alternatively, you can use `=`, but `<-` is widely preferred in the R community.

*** =pre_exercise_code
```{r}
 # no pec
```

*** =sample_code
```{r}
# Assign the value 7 to the variable my_oranges
# Print out the value of the variable my_oranges
```

*** =solution
```{r}
# Assign the value 7 to the variable my_oranges
my_oranges <- 7
# Print out the value of the variable my_oranges
my_oranges
```

*** =sct
```{r}
test_object("my_oranges", 
            undefined_msg = "Please make sure to define a variable `my_oranges`.",
            incorrect_msg = "Make sure that you assign the correct value to `my_oranges`.")
test_output_contains("my_oranges", incorrect_msg = "Have you explicitly told R to print out the `my_oranges` variable to the console?")
success_msg("Great! Continue to the next exercise!")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_5
## Variable assignment (3)

Every tasty fruit basket needs also apples, so you decide to add five apples to your oranges. As a data analyst, your reflex is to immediately create the variable `my_apples` and assign the value 5 to it. Next, you want to calculate how many pieces of fruit you have in total. Since you have given meaningful names to these values, you can now code this in a clear way: 

```
my_apples + my_oranges
```

*** =instructions
- Assign to `my_apples` the value 5.
- Add the variables `my_apples` and `my_oranges` and have R simply print the result.
- Assign the result of adding `my_apples` and `my_oranges` to a new variable `my_fruit`.

*** =hint
`my_fruit` is just the sum of `my_apples` and `my_oranges`. You can use the `+` operator to sum the two and `<-` to assign that value to the variable `my_fruit`.

*** =pre_exercise_code
`@pre_exercise_code`
```{r}
# no pec
```

*** =sample_code
```{r}
# Assign a value to the variables my_apples and my_oranges
my_oranges <- 7

# Add these two variables together
# Create the variable my_fruit
```
*** =solution
```{r}
# Assign a value to the variables my_apples and my_oranges
my_oranges <- 7
my_apples  <- 5
# Add these two variables together
my_apples + my_oranges
# Create the variable my_fruit
my_fruit <- my_apples + my_oranges
```

*** =sct
```{r}
test_object("my_apples", incorrect_msg = "Keep the line that assigns 5 to `my_apples`.")
test_object("my_oranges", incorrect_msg = "Keep the line that assigns 7 to `my_oranges`.")
test_output_contains("my_apples + my_oranges",
                     incorrect_msg = "Make sure to print out the result of adding `my_apples` and `my_oranges`. The code example in the description already gives away the answer to this instruction!")
msg <- "Have you used `my_fruit <- my_apples + my_oranges` to create the `my_fruit` variable?"
test_object("my_fruit", undefined_msg = msg, incorrect_msg = msg)
success_msg("Nice one! The great advantage of doing calculations with variables is reusability. If you just change `my_apples` to equal 12 instead of 5 and rerun the script, `my_fruit` will automatically update as well. Continue to the next exercise.")
```
--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_6
## Apples and oranges
Common knowledge tells you not to add apples and oranges. But hey, that is what you just did, no :-)? The `my_apples` and `my_oranges` variables both contained a number in the previous exercise. The `+` operator works with numeric variables in R. If you really tried to add "apples" and "oranges", and assigned a text value to the variable `my_oranges` (see the editor), you would be trying to assign the addition of a numeric and a character variable to the variable `my_fruit`. This is not possible.

`@instructions`
- Click 'Submit Answer' and read the error message. Make sure to understand why this did not work.
- Adjust the code so that R knows you have 6 oranges and thus a fruit basket with 11 pieces of fruit.

`@hint`
You have to assign the numeric value `6` to the `my_oranges` variable instead of the character value `"six"`. Note how the quotation marks are used to indicate that `"six"` is a character.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Assign a value to the variable my_apples
my_apples <- 5 
# Fix the assignment of my_oranges
my_oranges <- "six" 
# Create the variable my_fruit and print it out
my_fruit <- my_apples + my_oranges 
my_fruit
```

`@solution`
```{r}
# Assign a value to the variable my_apples
my_apples <- 5  
# Fix the assignment of my_oranges
my_oranges <- 6
# Create the variable my_fruit and print it out
my_fruit <- my_apples + my_oranges 
my_fruit
```

`@sct`
```{r}
test_error(incorrect_msg = "You can do this by setting the `my_oranges` variable to a numeric value, not a string!")
test_object("my_apples", incorrect_msg = "Make sure that `my_apples` still contains `5`.")
test_object("my_oranges", incorrect_msg = "Make sure that `my_oranges` is equal to `6`.")
test_object("my_fruit", incorrect_msg = "The value of `my_fruit` is not correct. It should be 11, the sum of `my_apples` and `my_oranges`.")
test_output_contains("my_fruit", incorrect_msg = "Don't remove the line that prints out `my_fruit`.")
success_msg("Awesome, keep up the good work! Continue to the next exercise.")
```

--- type:NormalExercise lang:r xp:0 skills:1 key:exercise0_7
## Basic data types in R

R works with numerous data types. Some of the most basic types to get started are:

- Decimal values like `4.5` are called **numerics**.
- Natural numbers like `4` are called **integers**. Integers are also numerics.
- Boolean values (`TRUE` or `FALSE`) are called **logical**.
- Text (or string) values are called **characters**.

Note how the quotation marks on the right indicate that "some text" is a character.

*** =instructions
Change the value of the:

- `my_numeric` variable to `42`.
- `my_character` variable to `"universe"`. Note that the quotation marks indicate that `"universe"` is a character.
- `my_logical` variable to `FALSE`.

Note that R is case sensitive!

*** =hint
Replace the values in the editor with the values that are provided in the exercise. For example: 
`my_numeric <- 42` assigns the value 42 to the variable `my_numeric`. 

*** =pre_exercise_code
```{r}
# no pec
```
*** =sample_code
```{r}
# Change my_numeric to be 42
my_numeric <- 42.5
# Change my_character to be "universe"
my_character <- "some text"
# Change my_logical to be FALSE
my_logical <- TRUE
```

*** =solution
```{r}
# Change my_numeric to be 42
my_numeric <- 42
# Change my_character to be "universe"
my_character <- "universe"
# Change my_logical to be FALSE
my_logical <- FALSE
```

*** =sct
```{r}
test_object("my_numeric", incorrect_msg = "Have you correctly changed the declaration of `my_numeric` so it contains the value 42?")
test_object("my_character", incorrect_msg = "Have you correctly changed `my_character` to `\"universe\"`? Don't forget the quotes!")
test_object("my_logical", incorrect_msg = "Have you correctly changed `my_logical` to `FALSE`? All letters of `FALSE` should be capitalized!")
success_msg("Great work! Continue to the next exercise.")
```

